---
title: "Compulsory Exercise 2: Title (give your project an informative title)"
author:
- "Einride B. Osland" \#1.
- "Magnus Engstrom" \#2.
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes: \usepackage{amsmath}
output:
  # html_document:
  #   toc: no
  #   toc_depth: '2'
  #   df_print: paged
  pdf_document:
    toc: no
    toc_depth: '2'
urlcolor: blue
abstract: "This is the place for your abstract (max 350 words)"
---

```{r setup, include=FALSE}
library(knitr)
# Feel free to change the setting as you see fit
knitr::opts_chunk$set(echo = TRUE,
                      tidy = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      strip.white = TRUE,
                      prompt = FALSE,
                      cache = TRUE,
                      size = "scriptsize",
                      fig.width = 4,
                      fig.height = 3,
                      fig.align = "center")

```

```{r, eval=TRUE, echo=FALSE}
library("knitr")
library("rmarkdown")
library(gbm)
library(caret)
```

<!--  Etc (load all packages needed). -->

## Introduction: Scope and purpose of your project

## Abstract (max. 350 words) (5 points)

In this project we are aiming to predict customer satisfaction of people who rent Airbnb apartments. This is going to help future customers to choose where to rent in order to have a better experience. It will also help Airbnb filter out apartments that tend to not give good satisfaction, and thus drive away customers, and instead recommend places that tend to make people return.


## Descriptive data analysis/statistics

## Methods

## Results and interpretation

## Summary



```{r}


Adam_w_data <- data.frame(read.csv("amsterdam_weekdays.csv"))
Adam_w_data <- subset(Adam_w_data, select = -X)
Adam_w_data$room_type <- as.factor(Adam_w_data$room_type)
Adam_w_data$room_shared <- as.factor(Adam_w_data$room_shared)
Adam_w_data$room_private <- as.factor(Adam_w_data$room_private)
Adam_w_data$host_is_superhost <- as.factor(Adam_w_data$host_is_superhost)



str(Adam_w_data)
```
### Principle component analysis


```{r}
df<-Adam_w_data


non_numeric_vars <- sapply(df, function(x) !is.numeric(x))

numeric_df <- df[, !non_numeric_vars]


pca_result <- prcomp(numeric_df, scale. = TRUE)

pca_components <- pca_result$x


variance_explained <- pca_result$sdev^2
```


```{r}



plot(pca_result)


plot(pca_components[, 1], pca_components[, 2], pch = 16, col = "blue", xlab = "PC1", ylab = "PC2")

```



```{r Gbm}
Adam_w_gbm <- gbm(formula = guest_satisfaction_overall ~., 
                  data = Adam_w_data,
                  distribution = "gaussian",
                  n.trees = 300,
                  shrinkage = 0.1,
                  interaction.depth = 5,
                  n.minobsinnode = 10,
                  bag.fraction = 1,
                  cv.folds = 5,
                  )

gbm.perf(Adam_w_gbm, method = "cv")
summary.gbm(Adam_w_gbm)

Adam_w_lm <- lm(formula = guest_satisfaction_overall ~., data = Adam_w_data)

K <- 5
folds <- 1:nrow(Adam_w_data)%%K

for (i in 0:(K-1)){
  train <- subset(data = Adam_w_data, folds != i)
}


summary(Adam_w_lm)





```


```

